Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-21 manuscript/manuscript.tex
Finished editing the conclusion
There was too much recap in the conclusion so I removed and summarized
some of it. I took out much of the algorithm and error analysis
description because it was repeated elsewhere. The conclusion should
give only a brief recap of the main results (in our case, the error
analysis). That's why I finished the conclusion with the results from
error analysis and a couple of sentences about why we can generalize
this to other density functions.

To overcome this challenge, we have developed a density-based discretization algorithm
that divides the tesseroid on the points at which the maximum variations of the density
function takes place.
The density-based discretization is performed before adaptive discretization as a type
of pre-processing step.
The number of divisions performed, and thus the accuracy of the computation, is
controlled by the delta ratio $\delta$.
This new algorithm is intended to minimise the error due to the inability of
the GLQ to produce precise approximations of highly variable density functions.

Because there is no direct relation between the values assigned to $D$ and
$\delta$ with the error of the computation, we had to empirically determine
the minimum value of $D$ and the maximum value of $\delta$ that produce an
acceptable accuracy for each gravitational field, while maintaining an
efficient computation.
The default values of $D$ and $\delta$ must produce accurate computations for
most continuous and smooth densities, that's why we have chosen two density
functions in order to establish them: a linear and an exponential one.

In the linear density case, the densty-based discretization algorithm is not
applied due to how it's build, so only a default value of $D$ had to be
determined.
By comparing the analytical solutions of the gravitational fields generated by
a spherical shell with the ones produced by the numerical model for different
values of $D$, we have obtained default values for the distance-size ratio of
1, 2 and 8 for the potential, its gradient and the Marussi tensor components,
respectively.
From these comparisons we can guarantee a precision of 0.1\% or greater on the
computations of the gravitational fields generated by any tesseroid with a
linear density.

The exponential density case is more complex, due to the fact that both $D$
and $\delta$ must be determined simultaneously.
We have explored regular gridded points in the $D$, $\delta$ space and
computed the percentage difference between the analytical solutions for a
spherical shell and the numerical model.
We had to take into account that the $D$ values cannot be lower than the ones
determined for the linear density case due to the fact that they must
guarantee acceptable accuracy for every continuous density function, including
the linear one.
From these comparisons we have concluded that the more efficient computations
take places when the $D$ values obtained in the linear density scenario are
used, i.e. the minimum ones.
This can be understood in terms of which kind of discretization produces more
subdivisions: a higher $D$ may produce more discretizations than a lower
$\delta$ because the first one divides the tesseroid in the radial, longitude
and latitude directions, while the latter only divides it in the radial
direction.
In conclusion, we found out that keeping the $D$ values as the ones obtained
for the linear density case, a $\delta = 0.2$ is needed for the computation of
the potential and its gradient components, while a $\delta = 0.01$ must be
used for the Marussi tensor components in order to ensure a 0.1\% precision.

In order to implement this new methodology we have developed a Python code
that carries out both discretization algorithms and the gravitational field
computation through the GLQ. It's freely available under the BSD 3-clause
open-source license and can be downloaded from
\todo[inline]{add repo url}

Finally, we have shown how this new method performs under a real data example:
we have computed the gravitational fields generated by the Neuqu\'en
sedimentary basin through a tesseroid model with an increasing exponential
density in depth on a regular grid at a computation height of 50km over the
reference ellipsoid.

>>>>

To overcome this challenge, we have developed a density-based discretization algorithm
that divides the tesseroid on the points at which the maximum variations of the density
function take place.
The density-based discretization is performed before adaptive discretization as a type
of pre-processing step.
The number of divisions performed, and thus the accuracy of the computation, is
controlled by the delta ratio $\delta$.
This new algorithm is intended to minimise the error due to the inability of
the GLQ to produce precise approximations of density functions with sharp variations.

Because there is no direct relation between the values assigned to $D$ and $\delta$ and
the error of the computation, we empirically determined the optimum values for these
parameters.
The values minimize the computational load while maintaining the numerical error below
0.1\% of an analytical solution.
The density functions used to establish the optimum values were a linear and an
exponential function.
The linear function represents the smoothest variation of density and does not require
density-based discretization at all.
We analyzed the error for exponential functions ranging from smooth to sharp variations
in depth to test the accuracy of the density-based discretization.
Through this analysis, we have obtained optimal values for the distance-size ratio of
1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
respectively.
These values are in agreement with previous results for a homogeneous density tesseroid
\citep{Uieda2016}.
The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
0.1\% precision.
A wide range of exponential functions were used to infer these results and they are
conservative estimates, chosen to prioritize accuracy over speed of computations.
Therefore, they can be safely generalized to other common depth-density relations, like
quadratic functions, without loss of accuracy.
